---
title: "Polychoric Correlation Analysis Demo"
format: html
editor: visual
---

# Polychoric Correlation Analysis in Evaluation Pipeline

This document demonstrates how to use the polychoric correlation analysis features added to the evaluation pipeline. Polychoric correlations are the appropriate method for analyzing relationships between ordinal variables (like Likert scale responses), as they provide unbiased estimates compared to Pearson correlations.

## Setup

```{r setup, message=FALSE}
# Load the evaluation pipeline
source("evaluation_pipeline.R")

# Load sample data (assuming you have evaluation data)
# data <- read.csv("your_evaluation_data.csv")  # Replace with your data file
```

## Key Features

### 1. Standalone Correlation Analysis

Calculate polychoric correlations between all ordinal questions:

```{r standalone-overall, eval=FALSE}
# Overall correlation analysis (all questions, all groups combined)
correlations <- analyze_correlations(data)

# Print summary
print_correlation_summary(correlations)

# Create visualization
plot_polychoric_correlations(correlations, method = "circle", 
                            title = "Overall Question Correlations")
```

### 2. Group-Specific Correlation Analysis

Analyze correlations within specific groups:

```{r group-specific, eval=FALSE}
# Instructor-level correlations
instructor_corr <- analyze_instructor_correlations(data)

# Subject-level correlations  
subject_corr <- analyze_subject_correlations(data)

# Custom grouping
semester_corr <- analyze_correlations(data, group_var = "semester")
```

### 3. Selective Question Analysis

Focus on specific questions of interest:

```{r selective-questions, eval=FALSE}
# Analyze correlations between specific questions
key_questions <- c("Q6_overall_instructor_rating", 
                   "Q7_course_quality",
                   "Q8_learning_objectives")

focused_corr <- analyze_correlations(data, question_stems = key_questions)

# Get summary of significant correlations
sig_summary <- summarize_significant_correlations(focused_corr, alpha = 0.05)
print(sig_summary)
```

### 4. Integration with Main Pipeline

Include correlation analysis as part of the main evaluation analysis:

```{r integrated-analysis, eval=FALSE}
# Full analysis with correlations included
full_analysis <- analyze_evaluations(
  data = data,
  group_var = "instructor",
  question_stem = "Q6_overall_instructor_rating",
  include_correlations = TRUE,
  correlation_questions = c("Q6_overall_instructor_rating", 
                           "Q7_course_quality", 
                           "Q8_learning_objectives"),
  min_responses = 50
)

# Access the correlation results
if (!is.null(full_analysis$correlations)) {
  print_correlation_summary(full_analysis$correlations)
  plot_polychoric_correlations(full_analysis$correlations)
}
```

## Understanding the Results

### Correlation Matrix

-   Values range from -1 to +1
-   Closer to ±1 indicates stronger relationship
-   Polychoric correlations assume underlying normal distribution

### Significance Testing

-   p-values test null hypothesis of zero correlation
-   Standard errors are typically larger than Pearson correlations
-   Confidence intervals provide range of plausible correlation values

### Visualization

-   Circle size represents correlation strength
-   Color intensity shows direction (positive/negative)
-   Significance stars indicate statistical significance:
    -   \*\*\* p \< 0.001
    -   \*\* p \< 0.01\
    -   
        -   p \< 0.05
    -   . p \< 0.1

## Advanced Options

### Custom Parameters

```{r advanced-options, eval=FALSE}
# Fine-tuned correlation analysis
custom_corr <- calculate_polychoric_correlations(
  data = data,
  question_stems = NULL,  # All questions
  group_var = "instructor",
  conf_level = 0.99,     # 99% confidence intervals
  min_responses = 25     # Lower threshold for inclusion
)
```

### Error Handling

The pipeline includes robust error handling: - Questions with insufficient responses are automatically excluded - Failed correlation calculations are reported but don't stop the analysis - Warning messages indicate which correlations couldn't be computed

### Performance Considerations

-   Polychoric correlations are computationally intensive
-   Consider using `min_responses` to filter questions
-   For large datasets, focus on key questions of interest

## Practical Applications

### 1. Construct Validity

Examine whether related questions correlate as expected:

```{r construct-validity, eval=FALSE}
# Teaching effectiveness questions should correlate
teaching_questions <- c("Q6_overall_instructor_rating",
                       "Q9_clear_explanations", 
                       "Q10_helpful_feedback")

teaching_corr <- analyze_correlations(data, question_stems = teaching_questions)
```

### 2. Question Redundancy

Identify highly correlated questions that might be redundant:

```{r redundancy-check, eval=FALSE}
# Find very high correlations (potential redundancy)
sig_corr <- summarize_significant_correlations(correlations)
highly_correlated <- sig_corr[abs(sig_corr$Correlation) > 0.8, ]
print(highly_correlated)
```

### 3. Factor Structure

Explore underlying factor structure in evaluation data:

```{r factor-structure, eval=FALSE}
# Correlation patterns can suggest factor structure
# High correlations within factors, lower between factors
all_corr <- analyze_correlations(data)
plot_polychoric_correlations(all_corr, method = "ellipse", 
                            title = "Factor Structure Analysis")
```

## Comparison with Pearson Correlations

The polychoric approach provides several advantages:

1.  **Unbiased estimates**: Pearson correlations on ordinal data are biased toward zero
2.  **Proper statistical inference**: Standard errors and confidence intervals are appropriate
3.  **Underlying assumptions**: Accounts for ordinal nature of the data

Note: Polychoric correlations will typically be larger in magnitude than Pearson correlations on the same ordinal data, but this represents the true underlying relationship rather than an artifact of the measurement scale.
